---
title: "Individual learner’s characteristic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Data processing
```{r,cache=TRUE,warning=FALSE,message=FALSE}
library(dplyr)
library(oulad)
library(ggplot2)
data(course)
data(assessment)
data(student)
data(vle)
data(student_assessment)
data(student_vle)
data(student_registration)
student_vle_unique<- student_vle %>% group_by(code_module,code_presentation,id_student,id_site,date) %>% summarise(sum_click=sum(sum_click))
rm(student_vle)
#Get student-level info across multiple datasets
tmaRatio<- assessment %>% filter(assessment_type!="Exam") %>% group_by(code_module,code_presentation) %>% summarise(tma=mean(assessment_type=="TMA"))

moduleStudentClick<-student_vle_unique %>% group_by(code_module,code_presentation,id_student) %>% summarise(meanDailyClick=sum(sum_click)/n_distinct(date))
mergedData<- merge(student,tmaRatio,by=c("code_module","code_presentation"))
mergedData<- merge(mergedData,moduleStudentClick,by=c("code_module","code_presentation","id_student"))

mergedData$code_module_category<-"Social sciences"
mergedData$code_module_category[mergedData$code_module %in% c("CCC","DDD","EEE","FFF")]<-"STEM"
#re-code final result and education level as ordered factor variables
uniq_levels<-unique(mergedData$highest_education)
mergedData$highest_education<-factor(mergedData$highest_education,levels=uniq_levels[c(5,2,3,1,4)],ordered=T)
mergedData$final_result<-factor(mergedData$final_result,levels=c("Withdrawn","Fail","Pass","Distinction"),ordered = T)

```

###Sex vs. final result###
Female students tend to do better in social science classes, male students do better in STEM classes. 
```{r,warning=FALSE,message=FALSE}
library(plotly)
library(tidyr) 
library(dplyr)
figure1 <- plot_ly(mergedData, x = ~final_result, color = ~gender) %>% add_histogram() %>% layout(title = 'Gender and Learning outcome',xaxis = list(type='category',title='Learning outcome'), yaxis=list (title='Total number of learners'),range = c(0,30000))
figure1
```

###Age vs. final result###
Younger students (<35) tend to do better, in both social science and STEM classes. 
```{r}
figure2 <- plot_ly(mergedData, x = ~final_result, color = ~age_band) %>% add_histogram() %>% layout(title = 'Age and Learning outcome',xaxis = list(type='category',title='Learning outcome'), yaxis=list (title='Total number of learners'),range = c(0,30000))
figure2
```

###Education level and final result###
It seems that better prior education levels are associated with better class outcome, especially STEM class outcome. In heat map, red means we observe more students with a particular combination of education level and final result than expected by chance.    
```{r}
figure3 <- plot_ly(mergedData, x = ~final_result, color = ~highest_education) %>% add_histogram() %>% layout(title = 'Education Level and Learning outcome',xaxis = list(type='category',title='Learning outcome'), yaxis=list (title='Total number of learners'),range = c(0,30000))
figure3
```

```{r,message=FALSE,warning=FALSE}
chisqResult<-chisq.test(mergedData$final_result,mergedData$highest_education)
chisqResult
ratioMat<-as.data.frame.matrix(chisqResult$observed/chisqResult$expected)
library(ComplexHeatmap)
library(circlize)
Heatmap(ratioMat,name="Ratio",cluster_rows = F,cluster_columns = F,col = colorRamp2(c(0, 1, 2.9), c("blue", "white", "red")))
```